Contracts for unions and intersections are hard for Nickel

### 2021-05-28 paper.tex
Pass on sec 4

\subsection*{[WIP] Union and intersection are not a silver bullet}

Before wrapping up this section, we'd like to show that maybe, union and intersection
contracts are not exactly what we would want them to be.
As stated throughout this work, programmers coming from dynamic languages
are accustomed to perform safety checks (data validation) with
great amounts of flexibility, and the whole power of their
language of choice at their disposition.

For instance, a Python programmer may have no problem in checking
at runtime, that a given variable is an integer if a given flag is
not set, as shown in Figure \ref{fig:python-assertion},
while a Haskell programmer might close her eyes and leave the room
instantly.

\begin{figure}[h]
\begin{lstlisting}[language=python]
result = 42 if flag else "Disabled"
...
assert ((not flag) or isinstance(result, int))
\end{lstlisting}
\caption{Asserting at runtime the type of a variable}
\label{fig:python-assertion}
\end{figure}

These expectations are not really met by how unions and intersection
contracts are supposed to work.
For example, consider the Nickel code shown on Figure
\ref{fig:intersection-distrbution}, at first sight, this code should
complete execution succesfully, returning \nickel{(1, 1)},
however, under both the Keil and Thiemann, as well as the
Williams, Morris, and Wadler works, this program will fail, with a violation
of the contract imposed to \nickel{g}.

@@

The reason this happens is that
there is a fundamental characterization of how intersection
contracts behave: the choice of which branch to choose is made exactly
when the intersection is opened up, and not delayed as much as possible.
This means, that the contract presented on Figure \ref{fig:intersection-distrbution},
\nickel{(Number -> Number -> Number) & (Number -> Boolean -> Number)},
will not behave the same way that
\nickel{(Number -> (Number -> Number) & (Boolean -> Number)},
when these two expressions should be logically equivalent.

@@

\subsection{Keil and Thiemann}
\label{sec:keil-thiemann}
\unsure{Subsection title?}
Starting with the Keil and Thiemann work \cite{KeilThiemannUnionIntersection},
they noticed that, since union contracts should share information between
different application contexts, while intersection contracts should behave
independently between each application context, the system becomes much
simpler if every contract gets transformed into a union of intersections
of other contracts, by using the usual De Morgan laws,
and then eagerly opening up unions, but postponing intersection
contracts as much as possible.
\todo{This assumes the difficulty has been previously explained}

@@

\subsection{Williams, Morris, and Wadler}
\label{sec:will-morr-wadl}
\unsure{Subsection title?}

@@

\subsection{Racket}
\label{sec:racket}

>>>>

\subsection*{Union and intersection are not a silver bullet}

Before wrapping up this section, we would like to show that maybe, union and intersection
contracts are not exactly what we would want them to be.
As stated throughout this work, programmers coming from dynamic languages
are accustomed to perform safety checks (data validation) with
great amounts of flexibility and the whole power of their
language of choice at their disposal.

For instance, a Python programmer may have no problem in checking
at runtime, that a given variable is an integer if a given flag is
not set, as shown in Figure \ref{fig:python-assertion};
while a Haskell programmer might close her eyes and leave the room
instantly.

\begin{figure}[h]
\begin{lstlisting}[language=python]
result = 42 if flag else "Disabled"
...
assert (
  (not flag) 
  or isinstance(result, int)
)
\end{lstlisting}
\caption{Asserting at runtime the type of a variable}
\label{fig:python-assertion}
\end{figure}

These expectations are not really met by how unions and intersections
contracts are supposed to work.
For example, consider the Nickel code shown on Figure
\ref{fig:intersection-distrbution}, at first sight, this code should
complete execution succesfully, returning \nickel{(1, 1)},
however, under both the Keil and Thiemann, as well as the
Williams, Morris, and Wadler works, this program will fail, with a violation
of the contract imposed to \nickel{g}.

@@

The reason this happens is that
there is a fundamental property of how intersection
contracts behave: the choice of which branch to choose is made exactly
when the intersection is opened up, and not delayed as much as possible,
as one would like.
This means, that the contract presented on Figure \ref{fig:intersection-distrbution},
\nickel{(Number -> Number -> Number) & (Number -> Boolean -> Number)},
will not behave the same way that
\nickel{(Number -> (Number -> Number) & (Boolean -> Number)},
when, with an intuitive view of how intersections should behave,
these two expressions should be logically equivalent.

It is worth noting, that intersection types do have this property,
since subtyping would allow to correctly type check a term like
\nickel{g 1} as both \nickel{Number -> Number} as well as
\nickel{Boolean -> Number}, and then deducing that \nickel{g 1}
can be given the type \nickel{(Number -> Number) & (Boolean -> Number)}.

@@

\subsection*{Keil and Thiemann}
\label{sec:keil-thiemann}
\unsure{Subsection title?}
Starting with the Keil and Thiemann work \cite{KeilThiemannUnionIntersection},
they noticed that, since union contracts should share information between
different application contexts, while intersection contracts should behave
independently between each application context, the system becomes much
simpler if every contract gets transformed into a union of intersections
of other contracts, by using the usual De Morgan laws,
and then eagerly opening up unions, but postponing intersection
contracts as much as possible.
\todo{This assumes the difficulty has been previously explained}

@@

\subsection*{Williams, Morris, and Wadler}
\label{sec:will-morr-wadl}
\unsure{Subsection title?}

@@

\subsection*{Racket}
\label{sec:racket}
